---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`scdown` is for annotation-after downstream analysis inside one dataset.

The package assumes that clustering and annotation are already done. The core
questions are:

- what clusters are present on the embedding?
- which genes define each cluster?
- do known marker panels support the annotation?
- which clusters are transcriptionally similar?
- which signatures or pathways are active?
- which ligand-receptor pairs are plausible?

## Example data

```{r}
library(scdown)

example_data <- scdown_example()
str(example_data, max.level = 1)
```

The example object contains:

- `cells`: one row per cell with `cell`, `sample`, and `annotation`
- `embedding`: two-dimensional coordinates
- `expr`: a sparse count matrix

## Create a `scdown` object

```{r}
obj <- scdown(
  example_data,
  annotation_col = "annotation",
  sample_col = "sample"
)

obj
```

## Run the minimal downstream workflow

```{r}
res <- run_core(
  obj,
  signatures = c("cytotoxic", "antigen_presentation")
)

names(res)
```

## Marker discovery

```{r}
head(res$markers$tables$top_markers)
```

## Marker-based annotation check

```{r}
head(res$annotation_check$tables$annotation_panel_scores)
```

## Average expression and similarity

```{r}
head(res$average_expression$tables$average_expression)
head(res$cluster_similarity$tables$cluster_similarity)
```

## Signature scoring

```{r}
head(res$signatures$tables$signature_score_by_annotation)
```

## Exploratory communication

```{r}
head(res$communication$tables$communication_pair_summary)
```

## Build a compact report

```{r}
report_path <- build_scdown_report(obj)
report_path
```

The report writes one HTML file plus CSV and PNG artifacts for each section.
