---
title: "Recommended Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recommended Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(scdown)
library(knitr)
```

`scdown` is designed for **annotation-after downstream work**.

That means:

- do QC, normalization, dimensionality reduction, integration, and clustering upstream
- do label assignment upstream
- use `scdown` to review the annotated object quickly
- hand off the final formal analyses to specialist tools when needed

## Generic positioning

```{r}
kable(recommended_pipeline("generic"))
```

## PBMC

Use `scdown` after standard PBMC preprocessing and annotation.

```{r}
kable(recommended_pipeline("pbmc"))
```

This is usually enough when the goal is:

- checking whether annotation looks sensible
- comparing known immune signatures
- sharing a compact report with collaborators

## Tumor immune

Tumor-infiltrating immune datasets are a strong fit for `scdown` as a review
layer, especially for:

- markers
- user-selected genes
- composition
- immune signatures
- exploratory communication

```{r}
kable(recommended_pipeline("tumor_immune"))
```

## Patient comparison

For donor-aware or patient-vs-patient comparisons, `scdown` is still useful,
but mainly as the layer that decides what should be promoted into formal models.

```{r}
kable(recommended_pipeline("patient_comparison"))
```

In this scenario the usual handoff is:

- `scdown` for review, prioritization, and compact reports
- `dreamlet` or `muscat` for differential state
- `miloR` for abundance-style shifts
- CellChat or another specialist framework when communication is a primary claim

## Minimal practical rule

If the question is:

- "What does this annotated dataset look like?"
- "Which cell types and signatures stand out?"
- "Can I quickly share a report?"

then `scdown` is a good home.

If the question is:

- "Is this the final model I will publish?"
- "Do I need covariates, random effects, or formal abundance modeling?"

then `scdown` should be the handoff point, not the final engine.
