---
title: "Recommended Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recommended Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(scdown)
library(knitr)
```

`scdown` is designed for **annotation-after downstream work**.

That means:

- do QC, normalization, dimensionality reduction, integration, and clustering upstream
- do label assignment upstream
- use `scdown` to review the annotated object quickly
- hand off the final formal analyses to specialist tools when needed

## Generic positioning

```{r}
kable(scope_table("generic"))
kable(recommended_pipeline("generic"))
```

The shortest rule is:

- use `scdown` for review, prioritization, and compact reporting
- hand off to specialist tools once an endpoint becomes a primary claim

## What stays inside `scdown`

`scdown` is a good home for:

- maps and annotation review
- quick marker review
- user-selected genes
- sample-level composition summaries
- signature summaries
- exploratory communication
- compact HTML reports

## What should hand off

```{r}
kable(scope_table("patient_comparison"))
```

When the question becomes formal, export a handoff object:

```{r eval=FALSE}
obj <- scdown(scdown_example())
export_for_handoff(obj, target = "muscat")
export_for_handoff(obj, target = "dreamlet")
export_for_handoff(obj, target = "miloR")
export_for_handoff(obj, target = "CellChat")
```

These exports do not replace the specialist package. They prepare the object you
would pass into the specialist workflow.

## PBMC

Use `scdown` after standard PBMC preprocessing and annotation.

```{r}
kable(recommended_pipeline("pbmc"))
```

This is usually enough when the goal is:

- checking whether annotation looks sensible
- comparing known immune signatures
- sharing a compact report with collaborators

## Tumor immune

Tumor-infiltrating immune datasets are a strong fit for `scdown` as a review
layer, especially for:

- markers
- user-selected genes
- composition
- immune signatures
- exploratory communication

```{r}
kable(recommended_pipeline("tumor_immune"))
```

## Patient comparison

For donor-aware or patient-vs-patient comparisons, `scdown` is still useful,
but mainly as the layer that decides what should be promoted into formal models.

```{r}
kable(recommended_pipeline("patient_comparison"))
```

In this scenario the usual handoff is:

- `scdown` for review, prioritization, and compact reports
- `dreamlet` or `muscat` for differential state
- `miloR` for abundance-style shifts
- CellChat or another specialist framework when communication is a primary claim

## Minimal practical rule

If the question is:

- "What does this annotated dataset look like?"
- "Which cell types and signatures stand out?"
- "Can I quickly share a report?"

then `scdown` is a good home.

If the question is:

- "Is this the final model I will publish?"
- "Do I need covariates, random effects, or formal abundance modeling?"

then `scdown` should be the handoff point, not the final engine.
